---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Rat.db

<!-- badges: start -->
[![R-CMD-check](https://github.com/datapumpernickel/Rat.db/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/datapumpernickel/Rat.db/actions/workflows/R-CMD-check.yaml)

[![codecov](https://codecov.io/gh/datapumpernickel/Rat.db/graph/badge.svg?token=J3JZD1O6AQ)](https://codecov.io/gh/datapumpernickel/Rat.db)
<!-- badges: end -->

The goal of Rat.db is to make it easy to query the arms transfer database (ATDB) from 
the [Stockholm International Peace Research Institute](https://www.sipri.org/databases/milex). 
They offer the most up-to-date, openly available data on arms transfers among states. 

## 🚧 Under construction 🚧 

* ⚠️ Please verify a few data points of downloaded data with the SIPRI website
to make sure that everything works, as this is currently under construction. 
* ⚠️ Please cite SIPRI when using the data and read the Sources and Methods 
section on their website concerning limitations linked below ⬇️. 

## ⚙️ Installation

You can install the development version of Rat.db from [GitHub](https://github.com/) with:

```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("datapumpernickel/Rat.db")
```

## 💲🔁️ Usage

### Retrieving Arms Transfer Data

The core function of `Rat.db` is `atdb_get_data()`, which retrieves arms transfer data from SIPRI's ATDB.

```{r, eval = FALSE}
library(Rat.db)
atdb_get_data(verbose = FALSE, start_year, end_year, cache = TRUE)
```

#### Parameters:
- `verbose`: Logical. If `TRUE`, prints additional details about the query process. Default is `FALSE`.
- `start_year`: Numeric. The start of the date range to query.
- `end_year`: Numeric. The end of the date range to query.
- `cache`: Logical. If `TRUE`, caches the response to avoid redundant requests. Default is `FALSE`.
- `tidy_cols` Logical. If TRUE, formats the column headers to a tidy format. Default is `TRUE`

#### Example Usage:

```{r, eval = FALSE}
# Retrieve data from 2000 to 2020, using cache
arms_data <- atdb_get_data(start_year = 2000, end_year = 2020)

# Retrieve data without using cache
arms_data_no_cache <- atdb_get_data(start_year = 2000, end_year = 2020, cache = TRUE)
```

## 🗄️ Caching Behavior

By default, `Rat.db` uses a disk-based caching system to store API responses. This prevents redundant queries to the SIPRI database, improving performance and reducing server load. However, keep in mind that this could prevent you from getting the most recent data, if the SIPRI data has been renewed, but you are still using cached data. This is why caching is by default disabled. 

#### Default Cache Settings:
- **Maximum Cache Size**: `1GB` (1024 * 1024^2 bytes)
- **Maximum Cache Age**: `1 year` (60\*60\*24\*365 seconds)
- **Maximum Cached Entries**: Unlimited (`Inf`)

#### Changing Cache Settings:
Users can override the default cache settings by setting environment variables before loading the package:

```{r, eval = FALSE}
Sys.setenv(RATDB_CACHE_MAX_SIZE = "500 * 1024^2")  # Set max cache size to 500MB
Sys.setenv(RATDB_CACHE_MAX_AGE = "60*60*24*180")  # Set max cache age to 180 days
Sys.setenv(RATDB_CACHE_MAX_N = "500")             # Limit cache to 500 entries
```

#### Clearing the Cache
If needed, users can manually prune the cache by running:

```{r, eval = FALSE}
cachem::cache_disk(dir = tools::R_user_dir('Rat.db', which = 'cache'))$prune()
```

This will remove expired entries, ensuring that outdated data is not used in queries. However, this also happens automatically with every request where `cache = TRUE`. 

You can also run `destroy()` to fully delete the cache. Note however, this requires a session restart to work with the package again afterwards. 

```{r, eval = FALSE}
cachem::cache_disk(dir = tools::R_user_dir('Rat.db', which = 'cache'))$destroy()
```


## 📝 Sources and Methods 

- Always verify a few data points manually on the SIPRI website.
- The dataset may have limitations; consult the SIPRI [**Sources and Methods**](https://www.sipri.org/databases/armstransfers/sources-and-methods) for details.
- TIV values are not comparable to actual prices of weapons, SIPRI writes: *SIPRI TIV figures do not represent sales prices for arms transfers. They should therefore not be directly compared with gross domestic product (GDP), military expenditure, sales values or the financial value of export licences in an attempt to measure the economic burden of arms imports or the economic benefits of exports. They are best used as the raw data for calculating trends in international arms transfers over periods of time, global percentages for suppliers and recipients, and percentages for the volume of transfers to or from particular states.*

## 🔒 Copyright 

The Stockholm International Peace Research Institute limits the copying and 
redistribution of its data to the following two use-cases: 

* the excerption of SIPRI copyrighted material for such purposes as criticism, 
comment, news reporting, teaching, scholarship or research in which the use 
is for non-commercial purposes
* the reproduction of less than 10 per cent of a published data set. 

Hence, this package does **not** contain any SIPRI data itself. It merely 
automatizes the access through the website, by making the corresponding POST 
request and cleaning the resulting xlsx file into tidy formats.

## 🤝 Citation 

Please make sure to cite SIPRI when using their data with: 

*Information from the Stockholm International Peace Research Institute (SIPRI) Arms Transfers Database*


You can get a citation for the package with: 


```{r citation, eval = F}
citation('Rat.db')
```

---

For more information, refer to the [SIPRI ATDB website](https://armstransfers.sipri.org/ArmsTransfer/).
